---
title: "2_price_mapping"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999) # explicitly disables scientific notation
```

```{r}
library(tidyverse)
library(stringr)
library(fuzzyjoin)
```

Importing Data from the following:
* NADAC website: https://dev.socrata.com/foundry/data.medicaid.gov/a4y5-998d downloaded as of 2020-02-06
Associated information:
Dataset Identifier: a4y5-998d
Total Rows: 7822360
Source Domain: data.medicaid.gov
Created: 11/7/2014, 10:54:00 AM
Last Updated: 2/4/2020, 10:57:49 AM
Category: Drug Pricing and Payment
Attribution: Centers for Medicare and Medicaid Services
License: Public Domain U.S. Government
Owner: Medicaid.gov
Endpoint Version: 2.1

* Crosswalk file is obtained from https://data.nber.org/data/ndc-to-labeler-code-product-code-crosswalk.html

```{r}
nadac <- read.csv(file.path("input","NADAC_20200206.csv"), header = TRUE)
# if order by date is needed:
# nadac <- nadac[order(as.Date(nadac$Effective_Date, format="%m/%d/%Y")),]
crosswalk <- read.csv(file.path("input","crosswalk.csv"), header = TRUE)
ndc_mfg <- readRDS("final.Rda")

# select for relevant columns from ndc_mfg
# ndc_mfg <- ndc_mfg %>% mutate(NDC_lookup = str_replace(PRODUCTNDC, "-", ""))
draft_ndc_mfg <- ndc_mfg %>% 
  select(PRODUCTNDC, LABELERNAME, Applicant, Applicant_Full_Name, NONPROPRIETARYNAME, DOSAGEFORMNAME, SUBSTANCENAME) %>%
  group_by(PRODUCTNDC, LABELERNAME, Applicant, Applicant_Full_Name, NONPROPRIETARYNAME, DOSAGEFORMNAME, SUBSTANCENAME) %>%
  summarize()
draft_ndc_mfg <- draft_ndc_mfg %>%
  mutate(MOLFORM = toupper(paste0(NONPROPRIETARYNAME, "_", DOSAGEFORMNAME)))

# convert crosswalk's NDC to string
crosswalk <- crosswalk %>% mutate(ndc = paste0(ndc))

# convert draft_nadac's NDC to string
draft_nadac <- nadac %>% 
  mutate(NDC = paste0(NDC)) %>%
  mutate(Effective_Date = as.Date(Effective_Date, format = "%m/%d/%Y"))

# subset nadac data into old (11 digits) or new (<10 digits)
draft_nadac_old <- draft_nadac %>% 
  filter(str_length(NDC) == 11)

# for new NDCs in nadac, pad with 0's to 10 digit, discard last 2 digits to obtain xxxx-xxxx NDC codes
draft_nadac_new <- draft_nadac %>% 
  filter(str_length(NDC) < 11) %>%
  mutate(NDC = str_pad(NDC, 10, pad = "0")) %>%
  mutate(NDC = paste0(substring(NDC, 1,4), "-", substring(NDC,5,8)))
# left_join nadac_new with mfg information
draft_nadac_mfg_new <- left_join(draft_nadac_new, draft_ndc_mfg, by = c("NDC" = "PRODUCTNDC"))

# for old NDCs in nadac, check through possible combinations and cross-check valid lblcode/prodcode with crosswalk
draft_nadac_old <- draft_nadac_old %>%
  mutate(NDC_lbl = ifelse(substring(NDC, 1, 4) %in% crosswalk$lblcode, substring(NDC, 1, 4), 
                          ifelse(substring(NDC,1,5) %in% crosswalk$lblcode, substring(NDC, 1, 5), "NA"))) %>% 
        # looks like there is no 4-4-2 config in this dataset
  mutate(NDC_prod = ifelse(str_length(NDC_lbl) == 4, substring(NDC, 5, 8),
                           ifelse((str_length(NDC_lbl) == 5) & (substring(NDC, 6, 6) == "0"), substring(NDC,7,9), 
                           ifelse((str_length(NDC_lbl) == 5) & (substring(NDC, 6, 6) != "0"), substring(NDC,6,9), "NA"))))
draft_nadac_old <- draft_nadac_old %>%
  mutate(NDC_lookup = paste0(NDC_lbl, "-", NDC_prod))

# left_join nadac_old with mfg information
draft_nadac_mfg_old <- left_join(draft_nadac_old, draft_ndc_mfg, by = c("NDC_lookup" = "PRODUCTNDC"))

# combine resultant nadac dataframe
draft_nadac_mfg_old <- draft_nadac_mfg_old %>%
  select(-one_of(c("NDC_lbl", "NDC_prod", "NDC_lookup")))
draft_nadac_mfg <- rbind(draft_nadac_mfg_old, draft_nadac_mfg_new)

```

Logic for parsing 11-digit NDC in NADAC:
Assuming it is 4-4-2 e.g. 2465-8030-410
Look up 2465: if 2465 is a valid lblcode then look up based on 4-4-2 config, if 2465 doesn't exist then NA
When we did this, turns out there is no 4-4-2 config in existing NADAC

Assuming it is 5-x-x e.g. 24658-0304-10
if the 6th digit is '0' then it is 5-3-2, but format to look up is 24658-304 (so drop the '0' at the 6th position)
if the 6th digit is not '0' then it is 5-4-1 and format to look up is 24658-xxxx (no need to drop any '0')
the overall result is still that the final 2 digits will be dropped. They constitute the package code, padded with leading 0's or not.

```{r}
nadac_mfg <- draft_nadac_mfg
saveRDS(nadac_mfg,file = "nadac_mfg.Rda")
```

